---
title: "p8105_hw5_xg2451"
output: github_document
---

## Problem 1
```{r}
# Function to check for duplicate birthdays
has_duplicates <- function(n) {
  birthdays <- sample(1:365, n, replace = TRUE)
  return(any(duplicated(birthdays)))
}

# Parameters
group_sizes <- 2:50
n_simulations <- 10000
probs <- numeric(length(group_sizes))

# Run simulations for each group size
for (i in seq_along(group_sizes)) {
  n <- group_sizes[i]
  results <- replicate(n_simulations, has_duplicates(n))
  probs[i] <- mean(results)
}

# Plot 
plot(group_sizes, probs, type = "l", col = "blue",
     xlab = "Group Size",
     ylab = "Probability of Shared Birthday",
     main = "Birthday Simulation")
abline(h = 0.5, col = "red", lty = 2)
legend("bottomright", legend = c("Probability", "50% Threshold"),
       col = c("blue", "red"), lty = c(1, 2))

```

**Comment:**

The plot illustrates how the probability that at least two people share a birthday increases with the group size. Notably, around a group size of 23, the probability crosses the 50% threshold. As the group size approaches 50, the probability nears 100%, indicating that it becomes almost certain for a duplicate birthday to occur in larger groups.

## Problem 2
```{r message = FALSE}
library(tidyverse)
library(broom)

# Set parameters
n <- 30
sigma <- 5
mus <- 0:6
n_simulations <- 5000

# Dataframe to store results
results <- data.frame()

# Loop over mu values
for (mu in mus) {
  # Simulate datasets
  for (i in 1:n_simulations) {
    # Generate data
    x <- rnorm(n, mean = mu, sd = sigma)
    # Perform t-test of H0: mu = 0
    test <- t.test(x, mu = 0)
    # Sample mean and p-value
    mu_hat <- mean(x)
    p_value <- test$p.value
    # Store results
    results <- rbind(results, data.frame(mu = mu, mu_hat = mu_hat, p_value = p_value))
  }
}

```

```{r}
# Compute power and average estimates for each mu
power_results <- results %>%
  group_by(mu) %>%
  summarize(
    power = mean(p_value < 0.05),
    avg_mu_hat = mean(mu_hat),
    avg_mu_hat_rejected = mean(mu_hat[p_value < 0.05], na.rm = TRUE)
  )

# Plot power vs. true mu
ggplot(power_results, aes(x = mu, y = power)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(
    title = "Power vs. True Mean (μ)",
    x = "True Mean (μ)",
    y = "Power (Proportion of Null Rejected)"
  ) +
  theme_minimal()
```

**Comment:**

The plot shows that the power of the test increases with the true mean (μ). When μ is 0, the power is approximately 5%, aligning with the significance level (α = 0.05). As μ increases from 0 to 6, the power increases from 5% to nearly 100%. This demonstrates that larger effect sizes increase the likelihood of correctly rejecting the false null hypothesis.

```{r}
# Plot average μ̂ vs. true μ for all samples
ggplot(power_results, aes(x = mu, y = avg_mu_hat)) +
  geom_line(color = "green") +
  geom_point(color = "green") +
  labs(
    title = "Average Estimated μ̂ vs. True μ",
    x = "True Mean (μ)",
    y = "Average Estimated μ̂"
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  theme_minimal()
```

**Comment:**

In this plot, the green line represents the average estimated μ̂ across all samples for each true μ. This line closely follows the dashed line representing the ideal case where the estimated μ̂ equals the true μ (y = x). This indicates that the sample mean is an unbiased estimator of the true mean across all samples.

```{r warning = FALSE}
# Overlay average μ̂ for rejected null samples
ggplot(power_results) +
  geom_line(aes(x = mu, y = avg_mu_hat), color = "green") +
  geom_point(aes(x = mu, y = avg_mu_hat), color = "green") +
  geom_line(aes(x = mu, y = avg_mu_hat_rejected), color = "red") +
  geom_point(aes(x = mu, y = avg_mu_hat_rejected), color = "red") +
  labs(
    title = "Average Estimated μ̂ vs. True μ",
    x = "True Mean (μ)",
    y = "Average Estimated μ̂"
  ) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
  theme_minimal()

```

**Comment:**

From the plot above, we can see when the effect size is small (<4 in this case), the sample average of μ̂ which the null is rejected is different from the true value of μ. When the effect size gets larger (>=4 in this case), the sample average of μ̂ which the null is rejected is approximately equal to the true value of μ. This is because the power is increasing as the effect size increases.














